%%%%%%%%%%%%%%%%%%%%%%%
% Comp 160, Fall 2019
% Homework 3
% Author: Vladimir Hugec
%%%%%%%%%%%%%%%%%%%%%%%

% This portion of the LaTeX document are configuration 
% You can see it as all the #includes in C++
\documentclass[12pt]{article}

\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{tikz}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}

\usepackage{titlesec}
\titleformat{\section}
{\normalfont\Large\bfseries}{Question~\thesection:}{1em}{}

\newlength{\toppush}
\setlength{\toppush}{2\headheight}
\addtolength{\toppush}{\headsep}

\def\subjnum{Comp 170}
\def\subjname{Computation Theory}

\def\doheading#1#2#3{\vfill\eject\vspace*{-\toppush}%
  \vbox{\hbox to\textwidth{{\bf} \subjnum: \subjname \hfil Vladimir Hugec}%
    \hbox to\textwidth{{\bf} Tufts University, Fall 2019 \hfil#3\strut}%
    \hrule}}


\newcommand{\htitle}[1]{\vspace*{1.25ex plus 1ex minus 0ex}%
\begin{center}
{\large\bf #1}
\end{center}} 


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN DOCUMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\doheading{2}{title}{Homework 3}

\section{}

$L_{neq} = \{w |$ w contains a different number of a’s, b’s, and c’s\}

\begin{proof}


Suppose $L_{neq}$ is a regular language. Let $p$ be the pumping length as defined by the Pumping Lemma. Let $w=a^j b^k c^l$, s.t. $j \neq k \neq l$. 

By pumping lemma, all words in the language longer than p can be pumped, $|w| \geq p$. Let $j= 2$, $k = 1$, and $l =3$ arbitrarily. And so $w = aabccc$. 

Divide the string, $w$ into parts $xyz$ s.t. $|xy| \leq p$ as follows: $x|y|z = aa|b|ccc$. Therefore by pumping lemma if $xyz \in L_{neq}$ then $xy^i z \in L_{neq}$ $\forall i \geq 0$. 

However, if we let $i=2$, then we have $xy^2 z = aa|bb|ccc$. However  $xy^2 z \notin L_{neq}$ since there are the same number of a's as well as b's.

There are two other possible ways to divide $w$ in $xyz$. $x|y|z = a|ab|ccc$ and $x|y|z = aa|bc|cc$. In both cases $xyz \notin L_{neq}$. In the former division, if we let $i=2$, then we have $xy^2 z = a|abab|ccc$, where the string contains an equal number of a's and c's. And in the latter division, if we let $i=2$ again, then we have $xy^2 z = aa|bcbc|cc$, where the string contains an equal number of a's and b's.

Therefore, no matter how the string $w$ is divided into $xyz$, there exist contradictions to the pumping lemma and therefore $L_{neq}$ is not regular. And since it is know that languages are closed under complement, the complement of $L_{neq}$ is also an irregular language

\end{proof}

\pagebreak

\section{}

$\indent L_{\geq} = \{w1w2 |$ $w2=b^n,n \geq 1$, and $w1 \in \Sigma^*$ contains $m \geq n $ a’s\} 

$L_{\leq} = \{w1w2 |$ $w2=b^n,n \geq 1$, and $w1 \in \Sigma^*$ contains $m \leq n $ a’s\} 

\subsection{A}

I could not figure out how to do a RegEx for this language. I felt like I've tried everything. Im not sure how on earth you limit the number of succeeding b's to be at most the number of a's present.

\subsection{B}

Suppose $L_{\leq}$ is a regular language. Let $p$ be the pumping length as defined by the Pumping Lemma. 

By pumping lemma, all words in the language longer than p can be pumped, $|w| \geq p$. Let $w = abaabb$ arbitrarily. 

Divide the string, $w$ into parts $xyz$ s.t. $|xy| \leq p$ as follows: $x|y|z = ab|a|abb$. Therefore by pumping lemma if $xyz \in L_{\leq}$ then $xy^i z \in L_{\leq}$ $\forall i \geq 0$. 

However, if we let $i=3$, then we have $xy^2 z = ab|aaa|abb$. However  $xy^2 z \notin L_{\leq}$ since the number of a's exceeds the number of succeeding b's.

There are other possible ways to divide $w$ in $xyz$. $x|y|z = aba|ab|b$. If we let $i=2$ here, we see that, once again, the number of a's exceeds the number of succeeding b's: $x|y^2|z = aba|abab|b$. Similarly, if we divide, $x|y|z = a|baa|bb$ and let $i=2$ again we see the same problem. No matter how this string gets divided there exist contradictions to the pumping lemma and therefore $L_{\leq}$ is not regular.


\pagebreak

\section{}

$L_{*} = \{w1*w2=w3 |$ w1, w2, and w3 are binary numbers, and w3 is the product of w1 and w2\}

\begin{proof}

Suppose $L_{*}$ is regular.  Let $p$ be the pumping length as defined by the Pumping Lemma. 

By pumping lemma, all words in the language longer than p can be pumped, $|w| \geq p$  Let $w = 1*10=10$ arbitrarily. 

Divide the string, $w$ into parts $xyz$ s.t. $|xy| \leq p$: $x|y|z = 1*|10|=10$. Therefore by pumping lemma if $xyz \in L_{*}$ then $xy^i z \in L_{*}$ $\forall i \geq 0$. 

However, if we simply let $i=2$, then we have $xy^2 z = 1*|1010| = 10$. However  $xy^2 z \notin L_{*}$ since this is not a valid binary multiplication.

Therefore, the language $L_{*}$ is not regular.

\end{proof}

\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

