%%%%%%%%%%%%%%%%%%%%%%%
% Comp 170, Fall 2019
% Homework 8
% Author: Vladimir Hugec
%%%%%%%%%%%%%%%%%%%%%%%

% This portion of the LaTeX document are configuration 
% You can see it as all the #includes in C++
\documentclass[12pt]{article}

\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{tikz}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}

\usepackage{titlesec}
\titleformat{\section}
{\normalfont\Large\bfseries}{Question~\thesection:}{1em}{}

\newlength{\toppush}
\setlength{\toppush}{2\headheight}
\addtolength{\toppush}{\headsep}

\def\subjnum{Comp 170}
\def\subjname{Computation Theory}

\def\doheading#1#2#3{\vfill\eject\vspace*{-\toppush}%
  \vbox{\hbox to\textwidth{{\bf} \subjnum: \subjname \hfil Vladimir Hugec}%
    \hbox to\textwidth{{\bf} Tufts University, Fall 2019 \hfil#3\strut}%
    \hrule}}


\newcommand{\htitle}[1]{\vspace*{1.25ex plus 1ex minus 0ex}%
\begin{center}
{\large\bf #1}
\end{center}} 


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN DOCUMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\doheading{2}{title}{Homework 8}

\section{}

Suppose that $END_{1}$ were decidable. Let $M$ be a Turing Machine that decides it. We can use $M$ to construct a Turing Machine that decides $A_{TM}$ as follows:

Input is $\langle X, w \rangle $ where $X$ is Turing Machine code and $w$ is a string.

Construct new Turing Machine $Y$ as follows:

$\bullet$ Input is a binary string, $s$.

$\bullet$ If the start of the string is a $0$ and  the end of the string there is a $1$, accept, else reject.

$\newline$

Run $M$ on input $\langle Y \rangle$, if $M$ accepts, accept, if $M$ rejects, reject.

If $Y$ accepts $s$, we know that $s$ is in the language and vice-versa if $Y$ rejects or loops. Therefore we can say that $M$ decides $A_{TM}$ which is a contradiction since we know that $A_{TM}$ is undecidable and therefore $END_{1}$ must be undecidable.

\pagebreak

\section{}

Suppose that $BLANK_{TM}$ were decidable. Let $M$ be a Turing Machine that decides it. We can use $M$ to construct a Turing Machine that decides $A_{TM}$ as follows:

Input is $\langle X, w \rangle $ where $X$ is Turing Machine code and $w$ is a string.

Construct new Turing Machine $Y$ as follows:

$\bullet$ Input is a binary string, $s$.

$\bullet$ At each character, erase it

$\bullet$ If $s$ ends, accept, else reject.

$\newline$

Run $M$ on input $\langle Y \rangle$, if $M$ accepts, accept, if $M$ rejects, reject.

If $Y$ accepts $s$, we know that $Y$ finished computation with a completely blank tape and vice-versa if $Y$ rejects or loops. Therefore we can say that $M$ decides $A_{TM}$ which is a contradiction since we know that $A_{TM}$ is undecidable and therefore $BLANK_{TM}$ must be undecidable.

\pagebreak

\section{}

\subsection{}

$\indent S \rightarrow XY | YX | \epsilon$

$X \rightarrow ZXZ | a | \epsilon$

$Y \rightarrow ZYZ | b \epsilon$

$Z \rightarrow a | b$

$\newline$ The grammar describes an arbitrary collection of a's and b's in a string $w$, $w \in \lbrace a,b \rbrace$*.

\subsection{}

The following $TM$, $X$, maps from $D$ to our grammer, $L$:

 Input is $\langle M, w \rangle$ where $M$ is a Turing Machine and $w$ is a string in our grammar above:
 
 $\bullet$ Replace each instance of the letter $0$ with $a$, and replace $1$ with $b$. "
 
 $\newline$
 
 What is left on the tape is now $\lbrace a,b \rbrace$*, so we can say $D$ maps to $L$. However since we know, $D$ is not context-free. Claim 1 must be false.
 
 \subsection{}
 
 The following $TM$, $Y$, maps from our grammer, $L$ to $D$:

 Input is $\langle M, w \rangle$ where $M$ is a Turing Machine and $w$ is a string in $D$:
 
 $\bullet$ Replace each instance of the letter $a$ with $0$, and replace $0$ with $1$. "
 
 $\newline$
 
 What is left on the tape is now $\lbrace 0,1 \rbrace$*, so we can say $L$ maps to $D$. However since we know, $L$ is context-free. Claim 2 must be false.
 
 
 






\enddocument
 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

